---
title: "Lecture 30"
author: "Albert Y. Kim"
date: "May 9, 2016"
output: ioslides_presentation
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Install these packages first
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(readr)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
param1 <- 100
param2 <- 4
obs <- 16
n <- 20
obs2 <- 80
n2 <- 100
```





## Today's Example

The parameter of interest is $p$: probability of flipping heads.  

* We believe beforehand that $p=0.5$.
* Our data says however $\widehat{p}=0.8$, suggesting $p=0.8$.





## Prior Distribution

The prior distribution reflects our **degree of belief** in different values of $p$:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
curve(dbeta(x, shape1 = param1, shape2 = param1), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Prior Distribution")
abline(v=0.5, lty=2)
```





## Likelihood Function

We flip a coin $n=`r n`$ times and observe $`r obs`$ heads. The likelihood function
along with the MLE $\widehat{p}=0.8$ of $p$ is:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
curve(dbinom(x = obs, size = n, prob = x), from=0, to=1, n=5000,
      xlab="p", ylab="Likelihood L(p|x)", main="Likelihood")
abline(v=obs/n, lty=2)
```





## Prior + Likelihood = Posterior 

We can then compute the posterior on the right:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
par(mfrow=c(2,3))
# First prior
curve(dbeta(x, shape1 = param1, shape2 = param1), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Prior Distribution")
abline(v=0.5, lty=2)
curve(dbinom(x = obs, size = n, prob = x), from=0, to=1, n=5000,
      xlab="p", ylab="Likelihood L(p|x)", main="Likelihood")
abline(v=obs/n, lty=2)
curve(dbeta(x, shape1 = param1+obs, shape2 = param1+n-obs), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Posterior Distribution")
abline(v=(param1+obs)/(param1+param1+n), lty=2)
```





## Two Comparisons

* Different priors centered at $p=0.5$, but one with
    + High prior certainty
    + Low prior certainty
* Different likelihood functions where $\widehat{p}=0.8$ but
    + $n=20$ and $16$ heads
    + $n=100$ and $80$ heads
    
    
    
    
    
## Comparison 1: Different Prior

Here, the red prior is less certain that $p=0.5$

```{r, echo=FALSE, message=FALSE, warning=FALSE}
curve(dbeta(x, shape1 = param1, shape2 = param1), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Prior Distribution")
abline(v=0.5, lty=2)
curve(dbeta(x, shape1 = param2, shape2 = param2), add=TRUE, col="red")
legend(
  "topright", legend = c("More Certain", "Less Certain"), bty="n",
  col=c("black", "red"), lty=c(1,1)
)
```





## Comparison 1: Different Prior

What does the posterior (with same likelihood) look like?

```{r, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(2,3))
# First prior
curve(dbeta(x, shape1 = param1, shape2 = param1), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Prior Distribution")
abline(v=0.5, lty=2)
curve(dbinom(x = obs, size = n, prob = x), from=0, to=1, n=5000,
      xlab="p", ylab="Likelihood L(p|x)", main="Likelihood")
abline(v=obs/n, lty=2)
curve(dbeta(x, shape1 = param1+obs, shape2 = param1+n-obs), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Posterior Distribution")
abline(v=(param1+obs)/(param1+param1+n), lty=2)
# Second prior
curve(dbeta(x, shape1 = param2, shape2 = param2), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Prior Distribution", col="red")
abline(v=0.5, lty=2, col="red")
```





## Comparison 1: Different Prior

The posterior in the bottom shifted to the right more.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(2,3))
# First prior
curve(dbeta(x, shape1 = param1, shape2 = param1), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Prior Distribution")
abline(v=0.5, lty=2)
curve(dbinom(x = obs, size = n, prob = x), from=0, to=1, n=5000,
      xlab="p", ylab="Likelihood L(p|x)", main="Likelihood")
abline(v=obs/n, lty=2)
curve(dbeta(x, shape1 = param1+obs, shape2 = param1+n-obs), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Posterior Distribution")
abline(v=(param1+obs)/(param1+param1+n), lty=2)
# Second prior
curve(dbeta(x, shape1 = param2, shape2 = param2), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Prior Distribution", col="red")
abline(v=0.5, lty=2, col="red")
curve(dbinom(x = obs, size = n, prob = x), from=0, to=1, n=5000,
      xlab="p", ylab="Likelihood L(p|x)", main="Likelihood")
abline(v=obs/n, lty=2)
curve(dbeta(x, shape1 = param2+obs, shape2 = param2+n-obs), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Posterior Distribution", col="red")
abline(v=(param2+obs)/(param2+param2+n), lty=2, col="red")
```







## Comparison 2: Different Likelihood

Now let the red curve reflect $n=`r n2`$ flips and $`r obs2`$ heads. Note that the 
likelihood function is more tight around 0.8.


```{r, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
curve(dbinom(x = obs, size = n, prob = x), from=0, to=1, n=5000,
      xlab="p", ylab="Likelihood L(p|x)", main="Likelihood")
abline(v=obs/n, lty=2)
curve(dbinom(x = obs2, size = n2, prob = x), from=0, to=1, n=5000,
      xlab="p", ylab="Likelihood L(p|x)", main="Likelihood", col="red")
abline(v=obs2/n2, lty=2, col="red")
```
    
    
    
    




## Comparison 2: Different Likelihood

What does the posterior (with same prior) look like?

```{r, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(2,3))
# First prior
curve(dbeta(x, shape1 = param1, shape2 = param1), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Prior Distribution")
abline(v=0.5, lty=2)
curve(dbinom(x = obs, size = n, prob = x), from=0, to=1, n=5000,
      xlab="p", ylab="Likelihood L(p|x)", main="Likelihood")
abline(v=obs/n, lty=2)
curve(dbeta(x, shape1 = param1+obs, shape2 = param1+n-obs), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Posterior Distribution")
abline(v=(param1+obs)/(param1+param1+n), lty=2)
# Second prior
frame()
curve(dbinom(x = obs2, size = n2, prob = x), from=0, to=1, n=5000,
      xlab="p", ylab="Likelihood L(p|x)", main="Likelihood", col="red")
abline(v=obs/n, lty=2, col="red")
```





## Comparison 2: Different Likelihood

The posterior in the bottom shifted to the right more.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(2,3))
# First prior
curve(dbeta(x, shape1 = param1, shape2 = param1), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Prior Distribution")
abline(v=0.5, lty=2)
curve(dbinom(x = obs, size = n, prob = x), from=0, to=1, n=5000,
      xlab="p", ylab="Likelihood L(p|x)", main="Likelihood")
abline(v=obs/n, lty=2)
curve(dbeta(x, shape1 = param1+obs, shape2 = param1+n-obs), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Posterior Distribution")
abline(v=(param1+obs)/(param1+param1+n), lty=2)
# Second prior
curve(dbeta(x, shape1 = param1, shape2 = param1), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Prior Distribution")
abline(v=0.5, lty=2)
curve(dbinom(x = obs2, size = n2, prob = x), from=0, to=1, n=5000,
      xlab="p", ylab="Likelihood L(p|x)", main="Likelihood", col="red")
abline(v=obs/n, lty=2, col="red")
curve(dbeta(x, shape1 = param2+obs2, shape2 = param2+n2-obs2), from=0, to=1, n=5000,
      xlab="p", ylab="density", main="Posterior Distribution", col="red")
abline(v=(param2+obs2)/(param2+param2+n2), lty=2, col="red")
```
