---
title: "Lecture 11"
author: "Albert Y. Kim"
date: "March 16, 2016"
output: ioslides_presentation
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Install these packages first
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(readr)

# Set seed for random number generator
set.seed(76)
```





## Today's Population

```{r, echo=FALSE}
# Note we are deleting 3 cases where there was no height recorded
profiles <- read_csv("profiles.csv") %>% 
  na.omit()
```

We consider 59K users who were

* members of OkCupid on 2012/06/26
* within 25 miles of SF
* online in the last year
* have at least one photo

Their public profiles were pulled on 2012/06/30






## Data

5 randomly chosen people:

```{r, echo=FALSE}
profiles %>% 
  sample_n(5) %>% 
  kable()
```





## Histogram

```{r, echo=FALSE, warning=FALSE}
ggplot(data=profiles, aes(x=height)) +
  geom_bar() +
  xlim(58, 80) +
  xlab("Height in inches") +
  ylab("Count") +
  theme_bw() +
  ggtitle("Heights of SF OkCupid Users") + 
  facet_wrap(~sex, nrow=2)
```





## Parameter of Interest

By what factor are men on average taller than women? i.e.

$$
\frac{\mu_M}{\mu_F}
$$

```{r, echo=FALSE}
# Note there are some missing values we are deleting
pop_means <- profiles %>% 
  group_by(sex) %>% 
  summarise(mean=mean(height))
pop_ratio <- pop_means$mean[2]/pop_means$mean[1]
```

Technically, we know the entire population, so we can compute this value exactly i.e. we don't need
statistics. 

Let's pretend however, that we only have a random sample of 10% of these users.

```{r, echo=FALSE}
# Note there are some missing values we are deleting
sample <- profiles %>% 
  sample_frac(.1) %>% 
  as.data.frame()
```





## Bootstrap Algorithm

```{r}
males <- subset(sample, select = height, subset = sex=="m", drop = T)
females <- subset(sample, select = height, subset = sex=="f", drop = T)

N <- 10^4
boot_ratio <- rep(0, N)

for(i in 1:N){
  boot_males <- sample(males, length(males), replace=TRUE)
  boot_females <- sample(females, length(females), replace=TRUE)
  
  boot_ratio[i] <- mean(boot_males)/mean(boot_females)
}
```





## Bootstrap Distribution

The point estimate is `r pop_ratio %>% round(3)`. The bootstrap distribution is:

```{r, echo=FALSE}
hist(boot_ratio, xlab="ratio", main="Ratio of Mean Male over Mean Female Height", probability = TRUE)
```





## Bootstrap Percentile Interval

We compute the percentile interval using the `quantile()` function

```{r}
quantile(boot_ratio, probs = c(0.025, 0.975))
```

The 95% bootstrap percentile-based CI is

$$
[1.079, 1.084]
$$





## Bootstrap Percentile Interval

We plot the CI with dashed lines:

```{r, echo=FALSE}
hist(boot_ratio, xlab="ratio", main="Ratio of Mean Male over Mean Female Height", probability = TRUE)
abline(v=quantile(boot_ratio, probs = c(0.025, 0.975)), lty=2, lwd=2)
```





## Bootstrap Percentile Interval

The true population ratio is `r round(pop_ratio,3)`. In this case, the CI captured it.

```{r, echo=FALSE}
hist(boot_ratio, xlab="ratio", main="Ratio of Mean Male over Mean Female Height", probability = TRUE)
abline(v=quantile(boot_ratio, probs = c(0.025, 0.975)), lty=2, lwd=2)
abline(v=pop_ratio, lwd=3, col="red")
```





## Interpretation of Confidence Intervals

If we were to repeat this whole procedure, say, 100 times:

1. Take a sample of 10% of the population
2. Compute the bootstrap CI

Then we expect 95 of the resulting CI's to capture the true value.





## Interpretation of Confidence Intervals

```{r, echo=FALSE, cache=TRUE}
N2 <- 100
CI <- data.frame(sim=1:N2, left=rep(0,N2), right=rep(0,N2), col=rep("black", N2))

for(j in 1:N2){
  sample <- profiles %>% sample_frac(.1) %>% as.data.frame()
  
  # Do bootstrap
  males <- subset(sample, select = height, subset = sex=="m", drop = T)
  females <- subset(sample, select = height, subset = sex=="f", drop = T)
  
  N <- 10^3
  boot_ratio <- rep(0, N)
  
  for(i in 1:N){
    boot_males <- sample(males, length(males), replace=TRUE)
    boot_females <- sample(females, length(females), replace=TRUE)
    
    boot_ratio[i] <- mean(boot_males)/mean(boot_females)
  }
  CI[j, 2:3] <- quantile(boot_ratio, probs = c(0.025, 0.975))
}

CI <- CI %>% 
  mutate(
    col = ifelse(pop_ratio > left & pop_ratio < right, "red", "black")
  ) %>% 
  gather("point", "value", 2:3)
```

The true ratio $\mu_M/\mu_F = 1.082$ is the dashed black line.

```{r, echo=FALSE}
ggplot(data=CI, aes(x=value, y=sim, group=sim, col=col)) + geom_line() +
  geom_vline(xintercept = pop_ratio, linetype = "longdash") +
  guides(colour=FALSE) +
  theme_bw() +
  xlab("Confidence Interval") + ylab("Simulation #") +
  ggtitle("100 Confidence Intervals Based on 100 Samples")
```





## In Practice

Wait! Why are we:

* Constructing more than one CI?
* Doing this if we already know the true population ratio $\mu_M/\mu_F = 1.082$ (dashed black line)

Much like our sampling distribution examples, this was a **theoretical/rhetorical**
exercise to demonstrate how CI's behave.






## In Practice

In practice, simply compute one CI based on one sample. This is our best
guess: **We are 95% "confident" the value lies in the CI.**






