---
title: "Lecture 08"
author: "Albert Y. Kim"
date: "March 7, 2016"
output: ioslides_presentation
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Install these packages first
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)

# Set seed for random number generator
set.seed(89)
```





## Bootstrap

Recall in HW01, Exercise 2.14, you simulated

* $n$ observations from a Normal$(0,1)$ random variable
* used a qqplot and a histogram to evaluate how normal the observations looked





## $n=15$

```{r, echo=FALSE}
x <- rnorm(15)
par(mfrow=c(1, 2))
qqnorm(x)
qqline(x)
hist(x, prob=TRUE)
# curve(dnorm(x, 0, 1), from=-4, to=4, n=100, add=TRUE)
```





## $n=100$

```{r, echo=FALSE}
x <- rnorm(100)
par(mfrow=c(1, 2))
qqnorm(x)
qqline(x)
hist(x, prob=TRUE)
# curve(dnorm(x, 0, 1), from=-4, to=4, n=100, add=TRUE)
```





## $n=1000$

```{r, echo=FALSE}
x <- rnorm(1000)
par(mfrow=c(1, 2))
qqnorm(x)
qqline(x)
hist(x, prob=TRUE)
# curve(dnorm(x, 0, 1), from=-4, to=4, n=100, add=TRUE)
```





## Moral of the Story

If the sample size $n$ is big, then the empirical distribution

```{r, echo=FALSE}
hist(x, prob=TRUE)
# curve(dnorm(x, 0, 1), from=-4, to=4, n=100, add=TRUE)
```





## Moral of the Story

well approximates the true population distribution.

```{r, echo=FALSE}
hist(x, prob=TRUE)
curve(dnorm(x, 0, 1), from=-4, to=4, n=100, add=TRUE, col="red", lwd=2)
legend("topright", legend="Pop Dist'n f(x)", col="red", lwd=2, bty="n")
```





## Bootstrap Motivation

We want to make statements about the **sampling distribution** of some statistic
based on samples of size $n$ from some population distribution.





## Bootstrap Motivation

```{r, echo=FALSE}
n <- 100000
sim <- c(
  rnorm(n, mean=-15, sd=3),
  rnorm(n, mean=10, sd=5),
  rnorm(n, mean=20, sd=10),
  rnorm(n, mean=-35, sd=10)  
)
dens.sim <- density(sim)
```

Typically we won't know the true population distribution...

```{r, echo=FALSE}
hist(sim, prob=TRUE, main="True Distribution", 
     ylim=c(0, 0.03), 
     xlab="x", ylab="f(x)", border="white")
lines(dens.sim, col="red", lwd=2)
```





## Bootstrap Motivation

... but we'll have a sample from it:

```{r, echo=FALSE}
hist(sim, prob=TRUE, main="Empirical Distribution", 
     ylim=c(0, 0.03), 
     xlab="x")
```





## Bootstrap Motivation

Then **resampling with replacement** from the empirical
distribution (left) approximates sampling from the true population distribution (right).


```{r, echo=FALSE}
par(mfrow=c(1,2))
hist(sim, prob=TRUE, main="Empirical Dist'n", 
     ylim=c(0, 0.03), 
     xlab="x", ylab="f(x)")
hist(sim, prob=TRUE, main="Population Dist'n", 
     ylim=c(0, 0.03), 
     xlab="x", ylab="f(x)", border = "white")
lines(dens.sim, col="red", lwd=2)
#legend("topright", legend="Pop Dist'n f(x)", col="red", lwd=2, bty="n")
```





## Example 1: We know result

Ex. where we know the result. Let the population distribution be 
Normal$(\mu=23, SD=7)$...

```{r, echo=FALSE}
curve(dnorm(x, mean=23, sd=7), from=0, to=45, ylab="Density", n=100,
      main="Population Distribution")
abline(v=23, lty=2)
```





## Example 1: We know result

```{r, echo=FALSE}
set.seed(79)
n <- 50
x <- rnorm(n, mean=23, sd=7)
```

from which you draw a sample of size $n=50$. We have $\overline{x}$ = `r mean(x) %>% round(3)` and:

```{r, echo=FALSE}
hist(x, xlim=c(0, 45), main="Empirical Distribution", prob=TRUE)
```




## Example 1: We know result

We **know** the sampling distribution of $\overline{X}$ is Normal$(\mu=23, SD=7/\sqrt{50}=0.99)$:

```{r, echo=FALSE}
curve(dnorm(x, mean=23, sd=7/sqrt(n)), from=18, to=27, ylab="Density", n=1000,
      main="Sampling Distribution", xlab="xbar", col="red")
abline(v=23, lty=2)
```





## Example 1: We know result

Run the algorithm:

```{r}
boot_sample_means <- rep(0, 10000)

for(i in 1:10000){
  boot_resample <- sample(x, replace = TRUE)
  boot_sample_means[i] <- mean(boot_resample)
}
```





## Example 1: We know result

The bootstrap distribution of $\overline{X}$...

```{r, echo=FALSE}
hist(boot_sample_means, xlim=c(18, 27), main="Bootstrap Distribution", xlab="xbar", prob=TRUE, breaks=50)
```





## Example 1: We know result

... is similar in shape and spread as the true sampling distribution but ...

```{r, echo=FALSE}
hist(boot_sample_means, xlim=c(18, 27), main="Bootstrap Distribution", xlab="xbar", prob=TRUE, breaks=50)
curve(dnorm(x, mean=23, sd=7/sqrt(n)), from=18, to=27, n=1000, add=TRUE, col="red")
legend("topright", legend="Sampling Dist'n", col="red", lty=1, bty="n")
```





## Example 1: We know result

... it is not centered at $\mu=23$, rather at $\overline{x}$ = `r mean(x) %>% round(3)`.

```{r, echo=FALSE}
hist(boot_sample_means, xlim=c(18, 27), main="Bootstrap Distribution", xlab="xbar", prob=TRUE, breaks=50)
curve(dnorm(x, mean=23, sd=7/sqrt(n)), from=18, to=27, n=1000, add=TRUE, col="red")
legend("topright", legend="Sampling Dist'n", col="red", lty=1, bty="n")
```






## Example 2: We don't know result

Now say we want to know the sampling distribution of a wacky statistic:

* $T(\vec{X})  = \frac{\max(\vec{X}) - \min(\vec{X})}{\text{median}(\vec{X})}$
* with $X_i \sim$ Normal$(\mu=27, \sigma=7)$ IID RV
* and $n=5000$






## Example 2: We don't know result

There is **no** result for what the sampling distribution of $T(\overline{X})$
is, or at least, it's hard to derive. So we create the bootstrap distribution:



## Example 2: We don't know result

```{r, echo=FALSE}
n <- 5000
x <- rnorm(n, mean=23, sd=7)
```

```{r}
boot_sample_WTF <- rep(0, 10000)

for(i in 1:10000){
  boot_resample <- sample(x, replace = TRUE)
  
  boot_sample_WTF[i] <- 
    (max(boot_resample) - min(boot_resample)) /
    median(boot_resample)
}
```





## Example 2: We don't know result

The the shape and spread of bootstrap distribution of $\overline{X}$ will be
similar to the shape and spread of the sampling distribution.

```{r, echo=FALSE}
hist(boot_sample_WTF, main="Bootstrap Distribution", xlab="xbar", prob=TRUE, breaks=50)
```





## Example 2: We don't know result

* it is "centered" at $T(\vec{x})$ = `r round((max(x) - min(x))/median(x), 3)`
* it has SD = `r round(sd(boot_sample_WTF), 3)`. i.e. the estimate of SE$[T(\vec{X})]$ is `r round(sd(boot_sample_WTF), 3)`.

```{r, echo=FALSE}
hist(boot_sample_WTF, main="Bootstrap Distribution", xlab="xbar", prob=TRUE, breaks=50)
abline(v=(max(x) - min(x))/median(x), col="red", lwd=2)
```





