---
title: "Lecture 06"
author: "Albert Y. Kim"
date: "March 2, 2016"
output: ioslides_presentation
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Install these packages first
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)

# Set seed for random number generator
set.seed(76)
```





## What are we doing here?

```{r, eval=FALSE}
students <- 
  c("Jigar", "James", "Davin", "Timothy", "Theodore", "Gurlyn", 
    "William", "Helena", "Kelsey", "Sarah", "Brenda", "Alvaro", 
    "Shrif", "Jules", "Henry", "Nimrod", "Graham", "Trisha", 
    "Vaasu", "Patrick", "Leo", "Yuchen")
sample(students, 3)
```





## Sampling Distribution by Simulation

Let $T(\vec{X}) = T(X_1, \ldots, X_n) = \max(\vec{X})$.

What is the sampling distribution of $T(\vec{X})$ when $n=12$? i.e. How does it behave?

We construct the sampling distribution **by simulation**:

* Let $X_i$ be Uniform$(0,1)$ RV for $i=1, \ldots, n$.
* Simulation 1000 values of $T(\vec{x})$ and histogram it.





## Sampling Distribution by Simulation

```{r, eval=FALSE}
sample_mins <- rep(0, 1000)
for(i in 1:1000){
  x_vec <- runif(n=12, min=0, max=1)
  sample_mins[i] <- max(x_vec)
}
hist(sample_mins)
```





## Sampling Distribution by Simulation

Count histogram based on 1000 simulations:

```{r, echo=FALSE}
sample_mins <- rep(0, 1000)
for(i in 1:1000){
  x_vec <- runif(n=12, min=0, max=1)
  sample_mins[i] <- max(x_vec)
}
hist(sample_mins, xlab="min", main="Sampling Distribution of max for n=12", breaks=seq(0.5, 1, by=0.025), xlim=c(0.6,1))
```





## Sampling Distribution by Simulation

Density histogram based on 1000 simulations:

```{r, echo=FALSE}
hist(sample_mins, xlab="min", main="Sampling Distribution of max for n=12", breaks=seq(0.5, 1, by=0.025), xlim=c(0.6,1), prob=TRUE)
```





## Sampling Distribution by Simulation

What do you think happens if we consider $T(\vec{X}) = \max(\vec{X})$ when $n=50$?





## Sampling Distribution by Simulation

Density histogram based on 1000 simulations:

```{r, echo=FALSE}
sample_mins <- rep(0, 1000)
for(i in 1:1000){
  x_vec <- runif(n=50)
  sample_mins[i] <- max(x_vec)
}
hist(sample_mins, xlab="min", main="Sampling Distribution of max for n=50", breaks=seq(0.5, 1, by=0.025), xlim=c(0.6,1), prob=TRUE)
```