---
title: "MLE for Normal Parameters"
author: "Albert Y. Kim"
date: "March 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## plotly Package

We install the [plotly](https://plot.ly/r/) R package

```{r, eval=FALSE}
devtools::install_github("ropensci/plotly")
```

and load it

```{r, message=FALSE, warning=FALSE}
library(plotly)
```


## Simulate Normal Data

We simulate a random sample of 1000 Normal$(\mu=0, \sigma=5)$ IID random variables.

```{r}
n <- 1000
x <- rnorm(n, mean=0, sd=5)
```



## MLE

We compute the MLE's $(\widehat{\mu}_{\text{MLE}}, \widehat{\sigma}_{\text{MLE}})$ for $(\mu, \sigma)$:

```{r}
xbar <- sum(x)/n
sigma_hat <- sqrt((1/n)*sum((x-xbar)^2))
c(xbar, sigma_hat)
```



## Log-Likelihood Function

The log-likelihood function can be computed, as described in the Lec12 notes, with the following function:

```{r}
f <- function(x, mu, sigma){
  sum(dnorm(x, mean=mu, sd=sigma, log = TRUE))
}
```

We plot the log-likelihood function for $(\mu, \sigma)$ given $\vec{x}$, saving the values in `log_lkhd` (code hidden).

```{r, echo=FALSE, cache=TRUE}
# Define 2D domain
mu <- seq(-1.5, 1.5, length=500)
sigma <- seq(4.3, 5.5, length=500)

# Compute log-likelihood
log_lkhd <- matrix(0, nrow=length(mu), ncol=length(sigma))
for(i in 1:nrow(log_lkhd))
  for(j in 1:ncol(log_lkhd))
    log_lkhd[i,j] <- f(x, mu=mu[i], sigma=sigma[j])
```

```{r}
plot_ly(z = log_lkhd, type = "surface", x=sigma, y=mu)
```



## Interpretation

The MLE $(\widehat{\mu}_{\text{MLE}}, \widehat{\sigma}_{\text{MLE}})$ are the parameter values that best explain/fit the observed data `x`.