---
title: "Lecture 09"
author: "Albert Y. Kim"
date: "March 11, 2016"
output: ioslides_presentation
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Install these packages first
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)

# Set seed for random number generator
set.seed(89)
```





## Example 1: We know result

Let the population distribution be Normal$(\mu=23, SD=7)$...

```{r, echo=FALSE}
curve(dnorm(x, mean=23, sd=7), from=0, to=45, ylab="Density", n=100,
      main="Population Distribution")
abline(v=23, lty=2)
```





## Example 1: We know result

```{r, echo=FALSE}
set.seed(79)
n <- 50
x <- rnorm(n, mean=23, sd=7)
```

from which you draw a sample of size $n=50$. We have $\overline{x}$ = `r mean(x) %>% round(3)` and:

```{r, echo=FALSE}
hist(x, xlim=c(0, 45), main="Empirical Distribution", prob=TRUE)
```





## Example 1: We know result

We **know** the sampling distribution of $\overline{X}$ is Normal$(\mu=23, SD=7/\sqrt{50}=0.99)$:

```{r, echo=FALSE}
curve(dnorm(x, mean=23, sd=7/sqrt(n)), from=18, to=27, ylab="Density", n=1000,
      main="Sampling Distribution", xlab="xbar", col="red")
abline(v=23, lty=2)
```





## Example 1: We know result

```{r, echo=FALSE}
boot_sample_means <- rep(0, 10000)

for(i in 1:10000){
  boot_resample <- sample(x, replace = TRUE)
  boot_sample_means[i] <- mean(boot_resample)
}
```

The bootstrap distribution of $\overline{X}$ is similar in shape and spread as the true sampling distribution, but it is not centered at $\mu=23$, rather at $\overline{x}$ = `r mean(x) %>% round(3)`.

```{r, echo=FALSE}
hist(boot_sample_means, xlim=c(18, 27), main="Bootstrap Distribution", xlab="xbar", prob=TRUE, breaks=50)
curve(dnorm(x, mean=23, sd=7/sqrt(n)), from=18, to=27, n=1000, add=TRUE, col="red")
legend("topright", legend="Sampling Dist'n", col="red", lty=1, bty="n")
```





## Standard Error

The **bootstrap standard error** (i.e. approximation) is:
```{r}
sd(boot_sample_means)
```

The **true standard error** is:
$$
\text{SE} = \frac{\sigma}{\sqrt{n}} = \frac{7}{\sqrt{50}} =0.99
$$





## Example 2: We don't know result

Now say we want to know the sampling distribution of a wacky statistic:

* $T(\vec{X})  = \frac{\max(\vec{X}) - \min(\vec{X})}{\text{median}(\vec{X})}$
* with $X_i \sim$ Normal$(\mu=27, \sigma=7)$ IID RV
* and $n=5000$






## Example 2: We don't know result

There is **no** result for what the sampling distribution of $T(\overline{X})$
is, or at least, it's hard to derive. So we create the bootstrap distribution:



## Example 2: We don't know result

```{r}
n <- 5000
x <- rnorm(n, mean=23, sd=7)

boot_sample_WTF <- rep(0, 10000)

for(i in 1:10000){
  boot_resample <- sample(x, replace = TRUE)
  
  boot_sample_WTF[i] <- 
    (max(boot_resample) - min(boot_resample)) / median(boot_resample)
}
```





## Example 2: We don't know result

The the shape and spread of bootstrap distribution of $\overline{X}$ will be
similar to that of the sampling distribution.

```{r, echo=FALSE}
hist(boot_sample_WTF, main="Bootstrap Distribution", xlab="xbar", prob=TRUE, breaks=50)
```





## Standard Error

The **bootstrap standard error** (i.e. approximation) is:
```{r}
sd(boot_sample_WTF)
```

The **true standard error** is: **?**





