---
title: "Lecture 03"
author: "Albert Y. Kim"
date: "February 22, 2016"
output: ioslides_presentation
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Install these packages first
library(dplyr)
library(readr)
library(ggplot2)
library(mosaic)
library(knitr)

# Set random number generator seed value explicitly so we can replicate our 
# simulations
set.seed(76)
```

```{r, echo=FALSE}
# Revisit grades datasest from Lec02
grades <- read_csv("../Lec02 ECDF and Permutation Test/grades.csv")
```




## Intro Stats Final Exam

Using the `grades.csv` dataset from `Lec02`, we study the
difference in mean final scores for

* those with an even # of letters in their last name vs.
* those with an odd #

i.e. the **parameter** of interest is $\mu_E - \mu_O$



## Intro Stats Final Exam

**Beware**: the solid line is the median, not the mean.

```{r, echo=FALSE}
ggplot(data=grades, aes(x=even, y=final)) + 
  geom_boxplot() +
  xlab("Even vs Odd") +
  ylab("Final Exam Score")
```



## Permutation Test

To describe the permutation test, we now consider

* only the first 6 rows
* and for illustrative purposes a new variable `row_index`

```{r, echo=FALSE}
# Only first 6 rows
grades_subset <- grades %>% 
  slice(1:6) %>% 
  select(-major) %>% 
  rename(even_vs_odd = even) %>% 
  mutate(
    row_index = 1:6,
    final = round(final, 3)
    ) 
```



## Intro Stats Final Exam

```{r, echo=FALSE}
grades_subset %>% kable()
```



## Observed Test Statistic

Rows `c(2, 3, 4, 6)` are the students with an even # of letters in their last
name.

To compute the **observed test statisitic** i.e. the observed difference in
sample means $\overline{x}_E - \overline{x}_O$ we run

```
index <- c(2, 3, 4, 6)

# mean of evens - mean of odds:
mean(final[index]) - mean(final[-index])
```



## The Crux

Under $H_0: \mu_E - \mu_O=0$, the two groups have equal mean. 

So under $H_0$ `even_vs_odd` is a meaningless label, hence we can
**permute/shuffle** it to no consequence.



## Observed Data

Originally, the evens are `c(2, 3, 4, 6)`

```{r, echo=FALSE}
grades_subset %>% 
  kable()
```



## Permuted Data

In this simulation, the evens are `c(1, 2, 3, 5)`

```{r, echo=FALSE}
grades_subset %>% 
  mutate(
    even_vs_odd = sample(even_vs_odd)
  ) %>% 
  kable()
```



## Permuted Data

In this simulation, the evens are `c(1, 2, 3, 5)`

```{r, echo=FALSE}
grades_subset %>% 
  mutate(
    even_vs_odd = sample(even_vs_odd)
  ) %>% 
  kable()
```



## Permuted Data

In this simulation, the evens are `c(1, 3, 4, 6)`

```{r, echo=FALSE}
grades_subset %>% 
  mutate(
    even_vs_odd = sample(even_vs_odd)
  ) %>% 
  kable()
```



## Permuted Data

In this simulation, the evens are `c(1, 3, 4, 5)`

```{r, echo=FALSE}
grades_subset %>% 
  mutate(
    even_vs_odd = sample(even_vs_odd)
  ) %>% 
  kable()
```



## Coding the Permutation

The permuting/shuffling of the `even_vs_odd` label is done by
randomly **resampling** which 4 of the 6 rows will be the evens:

```
index <- sample(6, size=4, replace = FALSE)

# Or equivalently:
row_indices <- 1:6
index <- sample(row_indices, size=4, replace = FALSE)
```

We sample **without replacement** b/c we don't want to resample any row more than once.

