---
title: "Lecture 29"
author: "Albert Y. Kim"
date: "May 6, 2016"
output: ioslides_presentation
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Install these packages first
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(readr)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
# Load data
profiles <- read_csv("../Lec22 More Hypothesis Testing/profiles.csv") %>% 
  tbl_df() %>% 
  mutate(sex = ifelse(is_female==1, "female", "male")) %>% 
  filter(height >=55 & height <= 80)
```





## OkCupid Data

Recall the OkCupid data set of 60K users in SF in 2012 from Lec22. We consider:

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
# Set seed for random number generator
set.seed(81)
profiles %>%  
  sample_n(5) %>% 
  select(-is_female) %>% 
  select(sex, height, has_wine) %>% 
  kable()
```





## Wine?

Does the user use the word "wine" in their essay responses?

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
has_wine_counts <- profiles %>% 
  group_by(has_wine) %>% 
  tally()

profiles %>%
  select(sex, has_wine) %>% 
  group_by(sex, has_wine) %>% 
  tally() %>% 
  inner_join(has_wine_counts, by="has_wine") %>% 
  mutate(prop=n.x/n.y) %>%
  select(sex, has_wine, prop) %>% 
  ggplot(data=., aes(x=has_wine, y=prop, fill=sex)) + 
  geom_bar(stat="identity", position="dodge") +
  labs(fill = "Sex", x='Has word "wine" in profile?', y="Proportion")
```





## Logistic Regression

```{r}
model <- glm(is_female ~ has_wine, data=profiles, family="binomial")
```

```{r, echo=FALSE}
kable(summary(model)$coef, digits=3)
```





## Fitted Probabilities

Since `has_wine` is binary, there are only two possible fitted probabilities:

```{r, echo=FALSE}
data_frame(prob = round(fitted(model),3)) %>%
  group_by(prob) %>%
  tally() %>%
  rename(
    `Fitted Prob` = prob,
    Number = n
  ) %>%
  kable()
```





## Height

```{r, echo=FALSE, message=FALSE, warning=FALSE}
profiles <- profiles %>% 
  filter(!is.na(height))

ggplot(data=profiles, aes(x=height)) +
  geom_histogram(aes(y=..density..), binwidth=1, boundary=55) + 
  facet_wrap(~sex, nrow=2) 
```





## Logistic Regression

```{r}
model <- glm(is_female ~ height, data=profiles, family="binomial")
```

```{r, echo=FALSE}
kable(summary(model)$coef, digits=3)
```





## Logistic Regression

To interpret the coefficient:

```{r}
exp(-0.662)
```





## Fitted Probabilities

For each height, we have a fitted probability of being female:

```{r, echo=FALSE}
profiles_sub <- profiles %>% 
  mutate(fitted = fitted(model)) %>% 
  distinct()

ggplot(data=profiles_sub, aes(x=height, y=fitted)) + 
  geom_point() +
  labs(x="Height", y="Fitted Probability of Being Female")
```



