---
title: "Lecture 17"
author: "Albert Y. Kim"
date: "April 6, 2016"
output: ioslides_presentation
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Install these packages first
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(readr)

# Set seed for random number generator
set.seed(76)
```



## t Distribution

Degrees of freedom = 1

```{r, echo=FALSE}
curve(dnorm(x, mean=0, sd=1), from=-4, to=4, n=1000, xlab="t", ylab="Density f(t)", 
      col="white", lty=2)
curve(dt(x, df=1), from=-4, to=4, n=1000, col="black", add=TRUE)
legend(
  "topright",
  bty="n",
  legend = c("t with df=1"),
  col = c("black"),
  lty = c(1)
)
```



## t Distribution

Degrees of freedom = 3

```{r, echo=FALSE}
curve(dnorm(x, mean=0, sd=1), from=-4, to=4, n=1000, xlab="t", ylab="Density f(t)", 
      col="white", lty=2)
curve(dt(x, df=3), from=-4, to=4, n=1000, col="red", add=TRUE)
curve(dt(x, df=1), from=-4, to=4, n=1000, col="black", add=TRUE)
legend(
  "topright",
  bty="n",
  legend = c("t with df=1", "t with df=3"),
  col = c("black", "red"),
  lty = c(1, 1)
)
```



## t Distribution

Degrees of freedom = 7

```{r, echo=FALSE}
curve(dnorm(x, mean=0, sd=1), from=-4, to=4, n=1000, xlab="t", ylab="Density f(t)", 
      col="white", lty=2)
curve(dt(x, df=7), from=-4, to=4, n=1000, col="green", add=TRUE)
curve(dt(x, df=3), from=-4, to=4, n=1000, col="red", add=TRUE)
curve(dt(x, df=1), from=-4, to=4, n=1000, col="black", add=TRUE)
legend(
  "topright",
  bty="n",
  legend = c("t with df=1", "t with df=3", "t with df=7"),
  col = c("black", "red", "green"),
  lty = c(1, 1, 1)
)
```



## t Distribution

At Degrees of freedom = $\infty$, the $t$ becomes a $Z \sim N(0,1)$

```{r, echo=FALSE}
curve(dnorm(x, mean=0, sd=1), from=-4, to=4, n=1000, xlab="t", ylab="Density f(t)", 
      col="blue", lty=2)
curve(dt(x, df=7), from=-4, to=4, n=1000, col="green", add=TRUE)
curve(dt(x, df=3), from=-4, to=4, n=1000, col="red", add=TRUE)
curve(dt(x, df=1), from=-4, to=4, n=1000, col="black", add=TRUE)
legend(
  "topright",
  bty="n",
  legend = c("t with df=1", "t with df=3", "t with df=7", "z"),
  col = c("black", "red", "green", "blue"),
  lty = c(1, 1, 1, 2)
)
```



## Example Page 175

```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4}
library(resampledata)
data("NCBirths2004")
girls <- subset(NCBirths2004$Weight, NCBirths2004$Gender == "Female")
```



## Example Page 175

Looks Normal!

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=4}
par(mfrow=c(1,2))
hist(girls, xlab="weight (in grams)", main="NC Girls Weights")
qqnorm(girls)
qqline(girls)
```



## Example Page 175

```{r, warning=FALSE, message=FALSE, fig.width=4, fig.height=4}
mean(girls)
sd(girls)
length(girls)
qt(0.975, df=521-1)
```



## Example Page 175

```{r}
t.test(girls, conf.level=0.95)
```



## Scenario 3

Starting in the "Assumption" section at the bottom of page 175, the book frames 
Scenario 3 as **a violation of the assumptions that the observations $X_i$ are
Normal**.

We will frame Scenario 3 as **CLT holding for observations irregardless of the
population distribution**.



