---
title: "Lecture 04"
author: "Albert Y. Kim"
date: "February 24, 2016"
output: ioslides_presentation
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Install these packages first
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
```

```{r, echo=FALSE, cache=TRUE}
# Load GSS2002 Data
GSS2002 <- source("https://raw.githubusercontent.com/Middlebury-Theory-of-Statistics/Non-Public/master/Datasets/GSS2002.R?token=AEXL-xTzrmcdfXdPzpSaJMasd1XUMMb1ks5W1j6ZwA%3D%3D")

GSS2002 <- GSS2002$value %>% 
  tbl_df() %>% 
  select(Education, DeathPenalty) %>% 
  filter(
    !is.na(Education),
    !is.na(DeathPenalty)
  )
```



## Chi-Square Test Statistics

We switch to tests involving Chi-Square Test Statistics.

**Motivating example:** Jurors are selected from the population of registered voters. How can we test if there is racial bias in juror selection?



## Jury Selection

Say you have a population of registered voters who break down as follows:

| Race              | White | Black | Hispanic | Other | Total |
|-------------------|-------|-------|----------|-------|-------|
| Registered Voters | 72%   | 7%    | 12%      | 9%    | 100%  |



## Jury Selection

Say you select $n=100$ jurors **at random** i.e. with no racial bias, we **expect** the following
jury representation

| Race              | White | Black | Hispanic | Other | Total |
|-------------------|-------|-------|----------|-------|-------|
| Registered Voters | 72%   | 7%    | 12%      | 9%    | 100%  |
| Jury Representation    | 72    | 7     | 12       | 9     | 100   |



## Jury Selection

Say you **observe** the following jury representation. Is there racial bias?

| Race              | White | Black | Hispanic | Other | Total |
|-------------------|-------|-------|----------|-------|-------|
| Registered Voters | 72%   | 7%    | 12%      | 9%    | 100%  |
| Jury Representation    | 76    | 7     | 9       | 8     | 100   |



## Jury Selection

Say you **observe** the following jury representation. Is there racial bias?

| Race              | White | Black | Hispanic | Other | Total |
|-------------------|-------|-------|----------|-------|-------|
| Registered Voters | 72%   | 7%    | 12%      | 9%    | 100%  |
| Jury Representation    | 0    | 100     | 0       | 0     | 100   |



## Today's Data

The General Social Survey is a major survey tracking American demographics, characteristics, and views on
social and cultural issues since the 1970's. 

Of the 2765 respondents in 2002, 1307 of them have values for both

* `Education`: Left HS, HS, Jr Col, Bachelors, or Graduate
* `DeathPenalty`: Favor, Oppose



## Today's Question

Are the two (categorical) variables `Education` and `DeathPenalty` **independent**?

If they are **dependent**, then the level of one variable will affect the level of the other.

i.e. People's support of the `DeathPenalty` will **depend** on your level of `Education`.



## Raw Data

Each of the 1307 rows represents one individual. Here are 6 randomly chosen rows:

```{r, echo=FALSE}
# Set random number generator seed value explicitly so we can replicate our 
# simulations
set.seed(2)
GSS2002_subset <- GSS2002 %>% 
  sample_n(6) %>% 
  kable()
GSS2002_subset
```



## Observed Contigency Table

The **observed** contingency table of `Education` and `DeathPenalty`:

```{r, echo=FALSE}
GSS2002 %>% 
  group_by(Education, DeathPenalty) %>% 
  tally() %>% 
  spread(DeathPenalty, n) %>% 
  ungroup() %>% 
  kable()
```



## Overall Proportion of Support

Column totals are below. So overall 898/1307 = 68.7% favor the death penalty.


```{r, echo=FALSE, warning=FALSE}
GSS2002 %>% 
  group_by(Education, DeathPenalty) %>% 
  tally() %>% 
  spread(DeathPenalty, n) %>% 
  ungroup() %>% 
  mutate(Total=Favor+Oppose) %>% 
  summarise(
    Favor=sum(Favor),
    Oppose=sum(Oppose),
    Total=sum(Total)
    ) %>% 
  mutate(Education = as.factor("Total")) %>% 
  select(Education, Favor, Oppose, Total) %>% 
  kable()
```




## Row Totals

We look at the row totals for each education level:

```{r, echo=FALSE}
GSS2002 %>% 
  group_by(Education, DeathPenalty) %>% 
  tally() %>% 
  spread(DeathPenalty, n) %>% 
  ungroup() %>% 
  mutate(Total=Favor+Oppose) %>% 
  kable()
```



## Back to Contingency Table

What would the contigency table look like if `Education` and `DeathPenalty` were
**independent**?

|Education | Favor| Oppose| Total|
|:---------|-----:|------:|-----:|
|Left HS   |   ? |     ?|   189|
|HS        |   ?|    ?|   711|
|Jr Col    |    ?|     ?|    87|
|Bachelors |   ?|     ?|   206|
|Graduate  |    ?|     ?|   114|



## Back to Contingency Table

What would the contigency table look like if `Education` and `DeathPenalty` were
**independent**?

|Education | Favor| Oppose| Total|
|:---------|-----:|------:|-----:|
|Left HS   |   189 x 0.687 = 129.8 |     ?|   189|
|HS        |   ?|    ?|   711|
|Jr Col    |    ?|     ?|    87|
|Bachelors |   ?|     ?|   206|
|Graduate  |    ?|     ?|   114|



## Back to Contingency Table

What would the contigency table look like if `Education` and `DeathPenalty` were
**independent**?

|Education | Favor| Oppose| Total|
|:---------|-----:|------:|-----:|
|Left HS   |   189 x 0.687 = 129.8 |     ?|   189|
|HS        |   711 x 0.687 = 488.5|    ?|   711|
|Jr Col    |    87 x 0.687 = 59.8|     ?|    87|
|Bachelors |   206 x 0.687 = 141.5|     ?|   206|
|Graduate  |    114 x 0.687 = 78.3|     ?|   114|



## Back to Contingency Table

And hence we can fill out the remaining column. This is the **expected** contigency table:

|Education | Favor| Oppose| Total|
|:---------|-----:|------:|-----:|
|Left HS   |   129.8 |     59.1|   189|
|HS        |   488.5|    222.5|   711|
|Jr Col    |   59.8|     27.2|    87|
|Bachelors |   141.5|     64.5|   206|
|Graduate  |   78.3|     35.7|   114|



## Chi-Square Statistic Based Tests

This an be done either using:

1. Computer: Permutation test for independence
1. Math: $\chi^2$ test for independence

We will also study the $\chi^2$ test for goodness-of-fit where

1. All parameters are known
1. Some parameters are estimated

