---
title: "Lecture 30"
author: "Albert Y. Kim"
date: "May 6, 2016"
output: ioslides_presentation
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(Lahman)
library(DT)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
career <- Batting %>% 
  tbl_df() %>% 
  # Remove pitchers and those with no AB
  anti_join(Pitching, by="playerID") %>% 
  filter(AB > 0) %>%
  # Create batters
  left_join(Master, by="playerID") %>% 
  unite(Name, nameFirst, nameLast, sep = " ") %>%
  # Sum over seasons. Note: some names are repeated
  group_by(playerID, Name) %>%
  summarize(H = sum(H), AB=sum(AB), Year=max(yearID)) %>%
  mutate(Avg = H / AB) %>% 
  ungroup() %>% 
  select(-playerID)
```





## Baseball Statistic

Most famous baseball statistic

$$
\text{Batting Avg} = \frac{\text{Hits}}{\text{At Bats}}
$$

which is a form of proportion $p$:

* $\text{Hits}$: success
* $\text{At Bats}$: Attempts at a Hit





## Baseball Data

Data on all `r nrow(career)` batters in Major League Baseball between 1871-2014:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
career %>% 
  mutate(Avg = round(Avg, 3)) %>% 
  select(-Year) %>% 
  datatable(options = list(pageLength=5, dom = 'tp'))
```





## Histogram of At Bats (Denominator)

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
ggplot(data=career, aes(x=AB)) + 
  geom_histogram(aes(y=..density..), boundary=0) +
  labs(x="At Bats", y="Count")
```

Quantiles:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
qtles <- quantile(career$AB, probs=c(0, 0.25, 0.5, 0.75, 1)) %>% round()
qtles
```





## Histogram of Batting Avgs

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
ggplot(data=career, aes(x=Avg)) + 
  geom_histogram(aes(y=..density..), boundary=0) +
  labs(x="Avg", y="Count")
```

Quantiles:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
quantile(career$Avg, probs=c(0, 0.25, 0.5, 0.75, 1)) %>% 
  round(3)
```





## Raw Highest Avgs

These are the batters with the best averages. Are they the best?

```{r, echo=FALSE, message=FALSE, warning=FALSE}
career %>% 
  arrange(desc(Avg)) %>%
  filter(Avg==1) %>% 
  mutate(Avg = round(Avg, 3)) %>% 
  select(-Year) %>% 
  datatable(options = list(pageLength=5, dom = 'tp'))
```





## Define Empirical Prior

```{r, echo=FALSE, warning=FALSE, message=FALSE}
career_filtered <- career %>%
    filter(AB >= 500)
overall_mean <- mean(career_filtered$Avg)
```

We use only those batters with at least 500 At Bats. The prior mean batting avg is
`r round(overall_mean, 3)`

```{r, echo=FALSE, warning=FALSE, message=FALSE}
prior_plot <- ggplot(data=career_filtered, aes(x=Avg)) + 
  geom_histogram(aes(y=..density..)) + 
  geom_vline(xintercept = overall_mean)
prior_plot
```





## Define Empirical Prior

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Find MLE's assuming Beta distribution
start_params <- list(shape1 = 1, shape2 = 10)
m <- MASS::fitdistr(career_filtered$Avg, dbeta, start = start_params)
alpha0 <- m$estimate[1]
beta0 <- m$estimate[2]

# Compute EBayes smoothed batting avgs.
career <- career %>%
    mutate(
      Shrunk_Avg = (H + alpha0) / (AB + alpha0 + beta0),
      AB_qtle = cut(career$AB, breaks=qtles, dig.lab=10, include.lowest=TRUE)
      )
```

Let's assume the prior is Beta$(\alpha, \beta)$. We find the MLE's $(\widehat{\alpha}, \widehat{\beta}) = 
(`r round(alpha0, 3)`, `r round(beta0, 3)`)$

```{r, echo=FALSE, warning=FALSE, message=FALSE}
prior_plot +
  stat_function(fun = dbeta, colour = "red", args=list(shape1=alpha0, shape2=beta0), size=1.5)
```





## Raw vs Shrunk Batting Avg

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Plots
career %>% 
  rename(
    `Raw Batting Avg` = Avg,
    `Shrunk Batting Avg` = Shrunk_Avg
  ) %>% 
  select(`Raw Batting Avg`, `Shrunk Batting Avg`) %>% 
  gather("type", "Avg", 1:2) %>% 
  ggplot(data=., aes(x=Avg, group=type)) +
  geom_histogram(boundary=0, aes(y=..density..), bins=50) +
  facet_wrap(~type, nrow=1) + 
  labs(fill="Batting Avg")
```





## Raw vs Shrunk Batting Avg

* Solid black horizontal line is overall batting avg
* Dashed diagonal line is $y=x$

```{r, echo=FALSE, warning=FALSE, message=FALSE}
base_plot <- 
  ggplot(data=career, aes(x=Avg, y=Shrunk_Avg, group=AB_qtle, col=AB_qtle)) +
  geom_hline(yintercept=overall_mean, size=0.25) +
  geom_abline(intercept=0, slope=1, linetype="dashed", size=0.25) +
  labs(x="Raw Batting Avg", y="Shrunk Batting Avg", color="At Bats")
base_plot + 
  geom_point()
```





## Raw vs Shrunk Batting Avg

```{r, echo=FALSE, warning=FALSE, message=FALSE}
base_plot +
  geom_point(size=1) +
  stat_smooth(method="lm", se=FALSE, fullrange = TRUE, size=0.25) +
  facet_wrap(~AB_qtle, nrow=2) +
  coord_fixed()
# ggsave("~/Desktop/Rplot.pdf", width=11*2/3, height=8.5*2/3)
```





## Highest Avgs

These are the best **adjusted** averages.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
career %>% 
  arrange(desc(Shrunk_Avg)) %>% 
  mutate(
    Avg = round(Avg, 3),
    Shrunk_Avg = round(Shrunk_Avg, 3)
    ) %>% 
  select(Name, H, AB, Shrunk_Avg) %>% 
  datatable(options = list(pageLength=5, dom = 'tp'))
```





## Highest Avgs

Note the year however:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
career %>% 
  arrange(desc(Shrunk_Avg)) %>% 
  mutate(
    Avg = round(Avg, 3),
    Shrunk_Avg = round(Shrunk_Avg, 3)
    ) %>% 
  select(Name, Year, H, AB, Shrunk_Avg) %>% 
  datatable(options = list(pageLength=5, dom = 'tp'))
```
